Fractional Types

### 2013-02-22 frac.tex
added proof of curry

\begin{multicols}{2}
  \dgm{eta_times2}

@@

In the following, we use {{bool}} as an abbreviation of {{1+1}} 
with {{true}} as an abbreviation for {{left ()}} and {{false}} as an 
abbreviation for {{right ()}}.

@@

>>>>

%% XXX: diagrams need unit wires(?)
\begin{multicols}{2}
  \dgm{eta_times2}

@@

In the following, we use {{bool}} as an abbreviation of {{1+1}} 
with {{true}} as an abbreviation for {{left ()}} and {{false}} as an 
abbreviation for {{right ()}}. In addition, for the sake of readability, the
circuit diagrams in this section will elide trivial unit wires, such as in
{{eta}} and {{eps}}.

@@

We can also derive currying (and dually, uncurrying) combinators. Observe that
the type of {{curry}} needs to be {{b1 * b2 :-* b3 <-> b1 :-* (b2 :-* b3)}},
which can be written in mathematical notation as {{b3/(b1 * b2) = (b3 * 1/b2) *
(1/b1)}}. This means {{curry}} can be written using {{recip}}, the
implementation of the mathematical identity {{1/(b1 * b2) = 1/b1 * 1/b2}}:

%subcode{opsem} include main
%! columnStyle = rcl
% recip &:& 1 / (b1 * b2) <-> 1 / b1 * 1 / b2
% recip &=& (uniti) (;) (uniti) (;) (assocl*) (;) ((eta* (*) eta*) (*) id)
%     &&    (;) (reorder (*) id) (;) (assocr*) (;) (id (*) swap*) (id (*) eps*)
%     &&    (;) swap* (;) (unite)

\noindent
where {{reorder}} is the obvious combinator of type {{b1 * (b2 * b3) * b4 <-> b1
* (b3 * b2) * b4}}. At the type level, the computation proceeds as follows:

%subcode{opsem} include main
%! columnstyle = rlr
%        & 1/(b1 * b2) & 
%     <->& 1 * 1/(b1 * b2) & uniti
%     <->& 1 * (1 * 1/(b1 * b2)) & uniti
%     <->& (1 * 1) * 1/(b1 * b2) & assocl*
%     <->& ((1/b1 * b1) * (1/b2 * b2)) * 1/(b1 * b2) &(eta*_{b1} (*) eta*_{b2}) (*) id
%     <->& ((1/b1 * 1/b2) * (b1 * b2)) * 1/(b1 * b2) & reoder (*) id
%     <->& (1/b1 * 1/b2) * ((b1 * b2) * 1/(b1 * b2)) & assocr*
%     <->& (1/b1 * 1/b2) * (1/(b1 * b2) * (b1 * b2)) & id * swap*
%     <->& (1/b1 * 1/b2) * 1 & id * eps*
%     <->& 1 * (1/b1 * 1/b2) & swap*
%     <->& (1/b1 * 1/b2) & unite

With {{recip}} out of the way, we can easily write {{curry}}:

%subcode{opsem} include main
%! columnStyle = rcl
% curry &:& b1 (x) b2 :-* b3 <-> b1 :-* (b2 :-* b3)
% curry &=& (id * recip) (;) (id * swap*) (;) assocl*
